---
title: "Summary output for EDA Project2"
author: "Tan Siu Ching"
date: "11/28/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Output Summary for EDA Project 2 

This is summary report about my output and answers to the questions for my EDA Project 2.  
The R code to the construction of the plot for Analysis are as follow:

### Questions 1
Have total emissions from PM2.5 decreased in the United States from 1999 to 2008? Using the base plotting system, make a plot showing the total PM2.5 emission from all sources for each of the years 1999, 2002, 2005, and 2008.


Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```
setwd("C:/Users/small/Documents/R_GitHub/EDA/Exploratory-Data-Analysis-EDA-Project-2")
library(dplyr)
Data_url <- "https://d396qusza40orc.cloudfront.net/exdata%2Fdata%2FNEI_data.zip" 
DaTZipfile <- "exdata_data_NEI_data.zip"
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
